---
title: "STAGES_plotting"
author: "Seán F"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r libs, include=FALSE}
library(ggplot2)
library(ggpubr)
library(dplyr)
library(circular)
library(purrr)
library(grid)
library(xfun)
library(purrr)
library(broom)
library(tidyr)
library(car)
library(broom)
library(lubridate)
library(stringr)
library(patchwork)
library(broom)
library(readr)
```
#load ata
```{r Data load}
#Load data from your directory, update this to work without my addresses
l5 <- read.csv("C:/Users/sefarrell/Downloads/sleep_timing_measures/l5_onset_offset_metrics_10min.csv")
cpd <- read_csv("C:/Users/sefarrell/Downloads/sleep_timing_measures/composite_phase_deviation_all_subjects.csv")
sjl <- read_csv("C:/Users/sefarrell/Downloads/sleep_timing_measures/social_jetlag_13window_from_outputs.csv")
std_dev_midsleep <- read_csv("C:/Users/sefarrell/Downloads/sleep_timing_measures/sleep_circular_stddev_HOURS.csv")
std_dev_duration <- read_csv("C:/Users/sefarrell/Downloads/sleep_timing_measures/sleep_circular_stddev_HOURS.csv")
sri <- read_csv("C:/Users/sefarrell/Downloads/sleep_timing_measures/SRI.csv")
is <- readxl::read_xlsx("C:/Users/sefarrell/Downloads/sleep_timing_measures/IS_summary_all_subjects.xlsx")

#Extract PSG subject IDs
target_subject_ids <- unique(psg_results_table$Subject_ID)

#Clean and filter data by subject_id
sri <- sri %>%
  mutate(subject_id = str_replace(File, "_binary\\.csv$", "")) %>%
  filter(subject_id %in% target_subject_ids)

is <- is %>%
  rename(subject_id = Subject_ID) %>%
  filter(subject_id %in% target_subject_ids)

sjl <- sjl %>% filter(subject_id %in% target_subject_ids)
std_dev_midsleep <- std_dev_midsleep %>% filter(subject_id %in% target_subject_ids)
std_dev_duration <- std_dev_duration %>% filter(subject_id %in% target_subject_ids)

if ("File" %in% names(cpd)) {
  cpd <- cpd %>%
    mutate(subject_id = str_replace(File, "_window.*$", "")) %>%
    filter(subject_id %in% target_subject_ids)
}

#Merge metrics into one df
cor_df_complete <- reduce(
  list(
    cpd %>% select(subject_id, CPD = mean_cpd),
    sjl %>% select(subject_id, SJL),
    std_dev_midsleep %>% select(subject_id, SD_Midsleep = midsleep_sd_hours),
    std_dev_duration %>% select(subject_id, SD_Duration = duration_sd_hours),
    sri %>% select(subject_id, SRI),
    is %>% select(subject_id, IS)
  ),
  ~left_join(.x, .y, by = "subject_id")
) %>% na.omit()

#Define analysis_df strictly from those in PSG and cor_df so there's no participants who overlap 
analysis_df <- psg_results_table %>%
  rename(subject_id = Subject_ID) %>%
  inner_join(cor_df_complete, by = "subject_id") %>%
  distinct(subject_id, .keep_all = TRUE)

#check all the appropriate people are in analysis df
ids_psg       <- unique(psg_results_table$Subject_ID)
ids_cor       <- unique(cor_df_complete$subject_id)
ids_analysis  <- unique(analysis_df$subject_id)

missing_from_cor       <- setdiff(ids_psg, ids_cor)
missing_from_analysis  <- setdiff(ids_psg, ids_analysis)
in_cor_not_in_analysis <- setdiff(ids_cor, ids_analysis)
in_analysis_not_in_cor <- setdiff(ids_analysis, ids_cor)

cat("Subjects in PSG but missing from cor_df_complete:\n")
print(missing_from_cor)

cat("\nSubjects in PSG but missing from analysis_df:\n")
print(missing_from_analysis)

cat("\nSubjects in cor_df_complete but missing from analysis_df:\n")
print(in_cor_not_in_analysis)

cat("\nSubjects in analysis_df but missing from cor_df_complete:\n")
print(in_analysis_not_in_cor)
#should be empty all going well
```
#Mean times and MEQ 
```{r Descriptives for table }
#vals are in decimal format so need to change to appropriate time vals
decimal_to_hm <- function(hours) {
  h <- floor(hours)
  m <- round((hours - h) * 60)
  if (m == 60) {
    h <- h + 1
    m <- 0
  }
  h <- h %% 24
  sprintf("%02d:%02d", h, m)
}
v_decimal_to_hm <- Vectorize(decimal_to_hm)
#use circular mean as before 
circular_mean_hours <- function(mins_or_hours) {
  vals_min <- ifelse(mean(mins_or_hours, na.rm = TRUE) < 24, mins_or_hours * 60, mins_or_hours)
  radians <- circular((vals_min %% 1440) / 1440 * 2 * pi, units = "radians")
  mean_rad <- mean.circular(radians)
  mean_hour <- as.numeric(mean_rad) * (1440 / (2 * pi)) %% 1440
  return(mean_hour / 60)
}
#get rid of duplicates
included_ids <- unique(analysis_df$subject_id)

sleep_timing_cleaned <- sleep_timing %>%
  filter(subject_id %in% included_ids) %>%
  mutate(
    onset_dt  = as.POSIXct(onset, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
    offset_dt = as.POSIXct(offset, format = "%Y-%m-%dT%H:%M:%SZ", tz = "UTC"),
    onset_hr  = as.numeric(format(onset_dt, "%H")) +
                as.numeric(format(onset_dt, "%M")) / 60 +
                as.numeric(format(onset_dt, "%S")) / 3600,
    offset_hr = as.numeric(format(offset_dt, "%H")) +
                as.numeric(format(offset_dt, "%M")) / 60 +
                as.numeric(format(offset_dt, "%S")) / 3600
  )
#set up new df with all timing summaries for group analysis
timing_summary <- sleep_timing_cleaned %>%
  summarise(
    onset_mean_hr     = circular_mean_hours(onset_hr),
    onset_sd_hr       = sd(onset_hr, na.rm = TRUE),
    offset_mean_hr    = circular_mean_hours(offset_hr),
    offset_sd_hr      = sd(offset_hr, na.rm = TRUE),
    duration_mean_min = mean(duration_mins, na.rm = TRUE),
    duration_sd_min   = sd(duration_mins, na.rm = TRUE)
  )
#get appropriate means
overall_summary_table <- tibble::tibble(
  Variable = c("Sleep Onset", "Sleep Offset", "Duration (min)"),
  Mean     = round(c(timing_summary$onset_mean_hr, timing_summary$offset_mean_hr, timing_summary$duration_mean_min), 2),
  SD       = round(c(timing_summary$onset_sd_hr, timing_summary$offset_sd_hr, timing_summary$duration_sd_min), 2),
  N        = c(
    n_distinct(sleep_timing_cleaned$subject_id[!is.na(sleep_timing_cleaned$onset_hr)]),
    n_distinct(sleep_timing_cleaned$subject_id[!is.na(sleep_timing_cleaned$offset_hr)]),
    n_distinct(sleep_timing_cleaned$subject_id[!is.na(sleep_timing_cleaned$duration_mins)])
  ),
  `HH:MM`  = c(
    v_decimal_to_hm(timing_summary$onset_mean_hr),
    v_decimal_to_hm(timing_summary$offset_mean_hr),
    sprintf("%02d:%02d",
            floor(timing_summary$duration_mean_min / 60),
            round(timing_summary$duration_mean_min %% 60))
  )
)
#scores for meq, used categorical defintions of eveningness/morningness
meq_tally <- raw_demographics %>%
  distinct(subject_code, .keep_all = TRUE) %>%
  filter(subject_code %in% included_ids) %>%
  mutate(
    MEQ_category = case_when(
      cir_0700 < 42 ~ "Evening",
      cir_0700 >= 42 & cir_0700 <= 58 ~ "Intermediate",
      cir_0700 > 58 ~ "Morning",
      TRUE ~ NA_character_
    ),
    MEQ_category = factor(MEQ_category, levels = c("Morning", "Intermediate", "Evening"))
  ) %>%
  count(MEQ_category, name = "n") %>%
  filter(!is.na(MEQ_category)) %>%
  arrange(MEQ_category)

print(overall_summary_table)
print(meq_tally)

```
#Sleep regularity measures summary
```{r Sleep reg descriptives}
#Summary of sleep regularity metrics for a table
reg_metrics <- tibble::tibble(
  Metric = c("SJL", "CPD", "SRI", "IS", "SD_Midsleep"),
  Mean = c(
    round(mean(sjl$SJL, na.rm = TRUE), 2),
    round(mean(cor_df_complete$CPD, na.rm = TRUE), 2),
    round(mean(cor_df_complete$SRI, na.rm = TRUE), 2),
    round(mean(cor_df_complete$IS, na.rm = TRUE), 2),
    round(mean(cor_df_complete$SD_Midsleep, na.rm = TRUE), 2)
  ),
  SD = c(
    round(sd(sjl$SJL, na.rm = TRUE), 2),
    round(sd(cor_df_complete$CPD, na.rm = TRUE), 2),
    round(sd(cor_df_complete$SRI, na.rm = TRUE), 2),
    round(sd(cor_df_complete$IS, na.rm = TRUE), 2),
    round(sd(cor_df_complete$SD_Midsleep, na.rm = TRUE), 2)
  ),
  N = c(
    sum(!is.na(sjl$SJL)),
    sum(!is.na(cor_df_complete$CPD)),
    sum(!is.na(cor_df_complete$SRI)),
    sum(!is.na(cor_df_complete$IS)),
    sum(!is.na(cor_df_complete$SD_Midsleep))
  )
)

if (requireNamespace("knitr", quietly = TRUE)) {
  cat("###Sleep Timing by Day Type\n")
  knitr::kable(
    diffs_for_report,
    caption = "Sleep Timing (Clock Time, HH:MM) and Mean Duration (Hours) by Day Type",
    col.names = c("Variable", "Free Days", "Work Days")
  )
  cat("\n###Sleep Regularity Metrics\n")
  knitr::kable(
    reg_metrics,
    caption = "Sleep Regularity Metrics (Mean, SD, N)"
  )
}
```
#Demographics
```{r Demograhics}
# load demographics file from appropriate place in dir
raw_demographics <- read.csv("C:/Users/sefarrell/Downloads/STAGES_questionnaire/Questionnaires/All_participant_demographics.csv")
#Remove duplicates from analysis_df by subject_id if needed, just a precaution as they should be already gone
analysis_df_unique <- analysis_df %>%
  distinct(subject_id, .keep_all = TRUE)
included_ids <- unique(analysis_df_unique$subject_id)

#Prepare demographics with missing as NA, do NOT add explicit "Not Reported" factor level, can change depending on what dmeographics are needed
demographics <- raw_demographics %>%
  filter(subject_code %in% included_ids) %>%
  distinct(subject_code, .keep_all = TRUE) %>%  #One row per subject
  transmute(
    subject_id = subject_code,
    sex = na_if(trimws(dem_0500), ""),
    sex = na_if(sex, " "),
    age = modified_dem_0110,
    BMI = dem_0800,
    race = dem_1000,
    race = ifelse(race == "", NA, race),
    smoker = current_cigarette_smoker,
    exercise = soclhx_0501,
    gad_score = gad_0800,
    phq9_score = phq_1000,
  ) %>%
  mutate(
    sex = factor(sex, levels = c("F", "M"), labels = c("Female", "Male")),
    race = factor(race, levels = 1:6, labels = c(
      "White/Caucasian",
      "Black or African American",
      "American Indian or Alaska Native",
      "Asian",
      "Pacific Islander",
      "Two races or some other race"
    )),
    smoker = factor(smoker, levels = c(0, 1), labels = c("No", "Yes")),
    exercise = factor(exercise, levels = c(0, 1), labels = c("No", "Yes"))
  )

#results table
demographics_table <- tibble(
  Variable = character(),
  N = integer(),
  `% or Mean (SD)` = character()
)

#exclude missing from N for continous vars
add_continuous <- function(var, label) {
  vals <- demographics[[var]]
  n_non_missing <- sum(!is.na(vals))
  m <- mean(vals, na.rm = TRUE)
  s <- sd(vals, na.rm = TRUE)
  demographics_table <<- add_row(demographics_table,
    Variable = label,
    N = n_non_missing,
    `% or Mean (SD)` = sprintf("%.2f (%.2f)", m, s)
  )
}

#each outcome on separate row for categorical for readability
add_categorical <- function(var, label) {
  vals <- demographics[[var]]
  vals <- vals[!is.na(vals)]  # exclude NA from counts
  tab <- table(vals)
  n <- sum(tab)
  
  #Add category headers
  demographics_table <<- add_row(demographics_table,
    Variable = label,
    N = n,
    `% or Mean (SD)` = ""
  )
  
  #each outcome as a separate row
  for (outcome in names(tab)) {
    count <- tab[outcome]
    percent <- sprintf("%.1f%%", 100 * count / n)
    demographics_table <<- add_row(demographics_table,
      Variable = paste0("  ", outcome),  # Indent subcategories
      N = as.integer(count),  # Show individual count in N column
      `% or Mean (SD)` = paste0("(", percent, ")")  # Just percentage in parentheses
    )
  }
}

#Add sample size row at the top
sample_size <- length(included_ids)
sample_size_row <- tibble(
  Variable = "Sample size",
  N = sample_size,
  `% or Mean (SD)` = ""
)
demographics_table <- bind_rows(sample_size_row, demographics_table)

#Build  table
add_continuous("age", "Age (years)")
add_categorical("sex", "Sex")
add_continuous("BMI", "BMI")
add_categorical("race", "Race")
add_categorical("smoker", "Current Smoker")
add_categorical("exercise", "Exercise Frequently")
add_continuous("gad_score", "GAD-7 Score")
add_continuous("phq9_score", "PHQ-9 Score")

#print with proper line breaks
print_demographics_table <- function(tbl) {
  bold_vars <- c(
    "Sample size", "Age (years)", "Sex", "BMI", "Race",
    "Current Smoker", "Exercise Frequency", "GAD-7 Score", "PHQ-9 Score"
  )
  
  for (i in seq_len(nrow(tbl))) {
    is_bold <- tbl$Variable[i] %in% bold_vars
    heading <- if (is_bold) paste0("\033[1m", tbl$Variable[i], "\033[0m") else tbl$Variable[i]
    
    cat(heading, "\n")
    cat("N:", tbl$N[i], "\n")
    cat(tbl$`% or Mean (SD)`[i], "\n\n")
  }
}

print_demographics_table(demographics_table)

#Save demographics_table, can render as .PDF later if needed
save(demographics_table, file = "C:/Users/sefarrell/Downloads/demographics_table.RData")
```
#Account for no. of people from each site
```{r Study site}
#Load data and extract locality (it's the subject ID prefix)
analysis_df_names <- analysis_df %>%
  mutate(site_code = substr(subject_id, 1, 4))

#Map site codes to locations
site_lookup <- tibble::tibble(
  site_code = c("BOGN", "GSBB", "GSDV", "GSLH", "GSSA", "GSSW", "MSNF", "MAYO", "MSQW", "MSTR", "STLK", "STNF", "MSMI", "MSTH"),
  location = c("South Carolina", "Pennsylvania", "Pennsylvania", "Pennsylvania", "Pennsylvania", "Pennsylvania",
               "Niagara Falls", "Minnesota", "Toronto", "Ontario", "Missouri", "California", "Ontario", "Toronto")
)

#Join and count
location_summary <- analysis_df_names %>%
  distinct(subject_id, .keep_all = TRUE) %>%
  left_join(site_lookup, by = "site_code") %>%
  count(location, name = "n_subjects") %>%
  mutate(
    percent = round(n_subjects / sum(n_subjects) * 100, 1),
    formatted = paste0(location, "\t", n_subjects, " (", percent, "%)")
  )

print(location_summary %>% select(location, formatted))
```
#Graph for distribution of SJL 
```{r SJL graph}
#plot Sjl acorss all inlcuded participants
ggplot(sjl, aes(x = SJL)) +
  geom_histogram(binwidth = 0.2, fill = "#0073C2FF", color = "white", alpha = 0.8, boundary = 0) +
  labs(
    title = "Distribution of Social Jetlag (SJL)",
    x = "SJL (hours)",
    y = "Number of Participants"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )
#Save as a  PNG
ggsave("C:/Users/sefarrell/Downloads/sjl_spread.png", width = 7, height = 4, dpi = 300)
```
#Plot via locality 
```{r Location}
#Just seperates SJL by locality 
sjl <- sjl %>% mutate(locality = substr(subject_id, 1, 4))
sleep_timing <- sleep_timing %>% mutate(locality = substr(subject_id, 1, 4))

ggplot(sjl, aes(x = locality, y = SJL, color = locality)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.18) +
  geom_jitter(width = 0.35, size = 2, alpha = 0.7) +
  labs(
    title = "Social Jetlag (SJL) by Locality",
    x = "Locality",
    y = "SJL (hours)"
  ) +
  theme_minimal(base_size = 18) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1, size = 16),
     plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(20, 20, 20, 20)
  )
ggsave("C:/Users/sefarrell/Downloads/sjl_by_locality_spread.png", width = 13, height = 5.5, dpi = 300)
```
#Plot concordance table 
```{r Concordance}
#summarize concordance between all sleep timing measures in analysis subset
included_ids <- unique(analysis_df$subject_id)
filtered_df <- cor_df_complete %>%
  filter(subject_id %in% included_ids)
metric_names <- c("SRI", "SD_Midsleep", "IS", "CPD", "SJL")

#Get all unique metric pairs (no repeats), should be okay but better safe than sorry
metric_pairs <- combn(metric_names, 2, simplify = FALSE)

#compute r and p for each pair of metrics
get_corr_stats <- function(pair, df) {
  x <- pair[1]
  y <- pair[2]
  test <- cor.test(df[[x]], df[[y]], method = "pearson")
  tibble(
    Comparison = paste(x, "vs.", y),
    r = round(test$estimate, 2),
    p_value = ifelse(test$p.value < 0.001, "<0.001", sprintf("%.3f", test$p.value))
  )
}

#Apply to all pairs and bind into table
cor_summary_table <- map_dfr(metric_pairs, get_corr_stats, df = filtered_df)

print(cor_summary_table)
write.csv(cor_summary_table,"C:/Users/sefarrell/Downloads/sleepregtable_final.csv")

```
#Sciencey graphs for concordance between metrics
```{r More plots}
#SRI vs. SD Midsleep
cor_A <- cor(cor_df_complete$SD_Midsleep, cor_df_complete$SRI, method = "pearson")

pA <- ggplot(cor_df_complete, aes(x = SD_Midsleep, y = SRI)) +
  geom_point(alpha = 0.5, size = 2, color = "#286090") +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  annotate(
    "text", 
    x = max(cor_df_complete$SD_Midsleep, na.rm = TRUE) - 0.1,
    y = max(cor_df_complete$SRI, na.rm = TRUE) - 2,
    label = sprintf("r = %.2f", cor_A),
    hjust = 1, vjust = 1, size = 6, fontface = "plain", color = "black"
  ) +
  labs(
    title = "Sleep Regularity Index vs. Midsleep Standard Deviation",
    x = "Midsleep Standard Deviation (hours)",
    y = "Sleep Regularity Index"
  ) +
  theme_bw(base_size = 16) +
  theme(plot.title = element_text(face = "plain", hjust = 0))

#SRI vs. IS
cor_B <- cor(cor_df_complete$SRI, cor_df_complete$IS, method = "pearson")

pB <- ggplot(cor_df_complete, aes(x = SRI, y = IS)) +
  geom_point(alpha = 0.5, size = 2, color = "#286090") +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  annotate(
    "text", 
    x = max(cor_df_complete$SRI, na.rm = TRUE) - 1,
    y = max(cor_df_complete$IS, na.rm = TRUE) - 0.01,
    label = sprintf("r = %.2f", cor_B),
    hjust = 1, vjust = 1, size = 6, fontface = "plain", color = "black"
  ) +
  labs(
    title = "Sleep Regularity Index vs. Interdaily Stability",
    x = "Sleep Regularity Index",
    y = "Interdaily Stability"
  ) +
  theme_bw(base_size = 16) +
  theme(plot.title = element_text(face = "plain", hjust = 0))

#Plot SJL by itself to visualize distribution 
median_sjl <- median(sjl$SJL, na.rm = TRUE)
pC <- ggplot(sjl, aes(x = SJL)) +
  geom_density(fill = "lightblue", color = "blue", alpha = 0.3) +
  geom_vline(xintercept = median_sjl, linetype = "dashed", color = "black", size = 1) +
  labs(
    title = "Distribution of Social Jetlag (SJL)",
    x = "Social Jetlag (hours)",
    y = "Density"
  ) +
  theme_bw(base_size = 16) +
  theme(plot.title = element_text(face = "plain", hjust = 0))

combined <- (pA / pB / pC) + 
  patchwork::plot_annotation(
    tag_levels = 'A', 
    tag_prefix = "", 
    tag_suffix = "."
  )

ggsave(
  "C:/Users/sefarrell/Downloads/sleep_metrics_panel.png",
  combined, width = 8, height = 15, dpi = 300, bg = "white"
)

```
#tab;es 
```{r tables}
#Select relevant columns
metrics <- cor_df_complete %>%
  dplyr::select(SJL, CPD, SRI, IS, SD_Midsleep)

#Summarise descriptives in long format for table
desc_df <- tibble(
  Metric = c("SJL", "CPD", "SRI", "IS", "SD_Midsleep"),
  Mean = c(
    mean(metrics$SJL, na.rm=TRUE),
    mean(metrics$CPD, na.rm=TRUE),
    mean(metrics$SRI, na.rm=TRUE),
    mean(metrics$IS, na.rm=TRUE),
    mean(metrics$SD_Midsleep, na.rm=TRUE)
  ),
  SD = c(
    sd(metrics$SJL, na.rm=TRUE),
    sd(metrics$CPD, na.rm=TRUE),
    sd(metrics$SRI, na.rm=TRUE),
    sd(metrics$IS, na.rm=TRUE),
    sd(metrics$SD_Midsleep, na.rm=TRUE)
  ),
  N = nrow(metrics)
)

desc_df %>%
  gt() %>%
  fmt_number(
    columns = c(Mean, SD),
    decimals = 2
  ) %>%
  tab_header(
    title = md("**Sleep Regularity Metrics**")
  ) %>%
  cols_label(
    Metric = md("**Metric**"),
    Mean = md("**Mean**"),
    SD = md("**SD**"),
    N = md("**N**")
  ) %>%
  opt_row_striping() %>%
  tab_options(
    table.width = pct(60),
    table.border.top.color = "teal",
    table.border.top.width = px(4),
    heading.background.color = "teal",
    column_labels.background.color = "teal",
    column_labels.font.weight = "bold",
    table_body.border.bottom.color = "black"
  )

```
#% models across 3 sleep stages with 3 levels of covariates
```{r % Models}
#Add BMI to demographics
demographics <- read.csv("C:/Users/sefarrell/Downloads/STAGES_questionnaire/Questionnaires/All_participant_demographics.csv") %>%
  dplyr::select(
    subject_id = subject_code,
    age         = modified_dem_0110,
    sex         = dem_0500,
    BMI         = dem_0800,
    gad_score   = gad_0800,
    phq9_score  = phq_1000
  ) %>%
  mutate(sex = factor(sex, levels = c("F", "M")))

#Select sleep timing metrics
timing_vars <- cor_df_complete %>%
  dplyr::select(subject_id, SD_Duration, SJL, SRI, IS)

#Select PSG percent and latency variables
psg_vars <- psg_results_table %>%
  rename(subject_id = Subject_ID) %>%
  dplyr::select(subject_id, N2_percent, N3_percent, REM_percent, REM_Lat_min, SOL)

#Merge datasets
analysis_df <- timing_vars %>%
  inner_join(psg_vars, by = "subject_id") %>%
  inner_join(demographics, by = "subject_id") %>%
  filter(complete.cases(.))
#=== Define model sets ===
dvs <- c("SRI", "SD_Duration", "SJL", "IS")

raw_predictors  <- "N2_percent + N3_percent + REM_percent + REM_Lat_min + SOL"
demo_predictors <- paste(raw_predictors, "+ age + sex + BMI")
full_predictors <- paste(demo_predictors, "+ phq9_score + gad_score")

model_list <- list(
  "Raw PSG only"     = raw_predictors,
  "Add demographics" = demo_predictors,
  "Add PHQ9 & GAD7"  = full_predictors
)

#Fit models and store results
model_results <- list()
vif_results   <- list()

for (dv in dvs) {
  for (label in names(model_list)) {
    fmla <- as.formula(paste(dv, "~", model_list[[label]]))
    model <- lm(fmla, data = analysis_df)

    #Coefficients
    tidy_df <- tidy(model, conf.int = TRUE) %>%
      mutate(outcome = dv, model_type = label) %>%
      select(outcome, model_type, term, estimate, conf.low, conf.high, p.value)

    #VIFs
    vif_df <- vif(model)
    vif_df <- tibble(
      outcome = dv,
      model_type = label,
      term = names(vif_df),
      VIF = as.numeric(vif_df)
    )

    model_results[[paste(dv, label)]] <- tidy_df
    vif_results[[paste(dv, label)]]   <- vif_df
  }
}

all_model_results <- bind_rows(model_results)
all_vif_results   <- bind_rows(vif_results)
cat("\n=== Model Estimates ===\n")
print(all_model_results)
cat("\n=== VIF Results ===\n")
print(all_vif_results)
```
#Secondary data analysis, load data 
```{r Second level}
cpd_path <- "C:/Users/sefarrell/Downloads/sleep_timing_measures/composite_phase_deviation_all_subjects.csv"
std_path <- "C:/Users/sefarrell/Downloads/sleep_timing_measures/sleep_circular_stddev_HOURS.csv"
#Read the CPD file
cpd <- read_csv(cpd_path)
print("CPD file:")
print(str(cpd))
#Read the onset/offset/midsleep SD file
std_dev_timing <- read_csv(std_path)
print("STD DEV file:")
print(str(std_dev_timing))
cpd_clean <- cpd %>%
  select(subject_id, mean_cpd)

std_clean <- std_dev_timing %>%
  select(subject_id, onset_sd_hours, offset_sd_hours, midsleep_sd_hours)
secondary_complete_df <- inner_join(cpd_clean, std_clean, by = "subject_id")
print(secondary_complete_df)
```
#Secondary models 3 stages, 3 levels 
```{r 2ndairy}
#reload demogrpahcis, just out this here in case only second one needs to be run/vice versa
demographics <- read.csv("C:/Users/sefarrell/Downloads/STAGES_questionnaire/Questionnaires/All_participant_demographics.csv") %>%
  dplyr::select(
    subject_id = subject_code,
    age         = modified_dem_0110,
    sex         = dem_0500,
    BMI         = dem_0800,
    gad_score   = gad_0800,
    phq9_score  = phq_1000
  ) %>%
  mutate(sex = factor(sex, levels = c("F", "M")))

#Select PSG architecture variables
psg_vars <- psg_results_table %>%
  rename(subject_id = Subject_ID) %>%
  dplyr::select(subject_id, N2_percent, N3_percent, REM_percent, REM_Lat_min, SOL)

#Select DVs from secondary_complete_df
secondary_dvs <- secondary_complete_df %>%
  dplyr::select(subject_id, onset_sd_hours, offset_sd_hours, midsleep_sd_hours, mean_cpd)

#Merge data
analysis_df_secondary <- secondary_dvs %>%
  inner_join(psg_vars, by = "subject_id") %>%
  inner_join(demographics, by = "subject_id") %>%
  filter(complete.cases(.))

#Define outcomes and model formulas
dvs_secondary <- c("onset_sd_hours", "offset_sd_hours", "midsleep_sd_hours", "mean_cpd")

model_list_secondary <- list(
  "Raw PSG only"     = "N2_percent + N3_percent + REM_percent + REM_Lat_min + SOL",
  "Add demographics" = "N2_percent + N3_percent + REM_percent + REM_Lat_min + SOL + age + sex + BMI",
  "Add PHQ9 & GAD7"  = "N2_percent + N3_percent + REM_percent + REM_Lat_min + SOL + age + sex + BMI + phq9_score + gad_score"
)

#Run models and store results
model_results_secondary <- list()
vif_results_secondary   <- list()

for (dv in dvs_secondary) {
  for (label in names(model_list_secondary)) {
    fmla <- as.formula(paste(dv, "~", model_list_secondary[[label]]))
    model <- lm(fmla, data = analysis_df_secondary)

    tidy_df <- broom::tidy(model, conf.int = TRUE) %>%
      mutate(outcome = dv, model_type = label) %>%
      select(outcome, model_type, term, estimate, conf.low, conf.high, p.value)

    vif_df <- car::vif(model)
    vif_df <- tibble(
      outcome = dv,
      model_type = label,
      term = names(vif_df),
      VIF = as.numeric(vif_df)
    )

    model_results_secondary[[paste(dv, label)]] <- tidy_df
    vif_results_secondary[[paste(dv, label)]]   <- vif_df
  }
}

all_model_results_secondary <- bind_rows(model_results_secondary)
all_vif_results_secondary   <- bind_rows(vif_results_secondary)

save(
  all_model_results,
  all_model_results_secondary,
  model_list,
  model_list_secondary,
  analysis_df,
  analysis_df_secondary,
  file = "C:/Users/sefarrell/Downloads/analysis_objects.RData"
)

```